Online {\IterComp}

### 2017-09-12 cgo18mscr.tex
adds formal description of instrumentation

\section{The Work-Efficiency Performance Metric} \label{sec:metric}

@@

By measuring the amount of \textit{computational work} done per unit of time,
it reduces the impact of input-dependent aspects and focus instead on the
efficiency of the optimized program.
For this metric, the main challenge is to precisely define what represents
\textit{work}.
In the context of experimental algorithmics, previous work has suggested using
\textit{block frequency} as a metric for empirical computational
complexity~\cite{goldsmith07}.
%Block frequency is a relative metric that represents the number of times a basic
%block executes~\citep{ball94,ball96}.

However, in the context of comparing different optimization sequences,
although block frequency would be able to capture aspects of optimizations that
simplify the control-flow graph (CFG), measuring work at the basic block
resolution would not capture effects of optimisations at the instruction level.
Because of that, in this thesis, we extend the idea of using basic block frequency to measure computational work by also considering the computational cost of each basic block.
The computational cost of a basic block is given by weighting the instructions that it contains.

@@

@@

>>>>

\section{The Work Efficiency Metric} \label{sec:metric}

@@

By measuring the amount of \textit{computational work} done per unit of time,
it reduces the impact of input-dependent aspects and focus instead on the
efficiency of the optimized program.
For this metric, the main challenge is to precisely define what represents
\textit{work}.
In the context of experimental algorithmics, previous works have suggested using
\textit{block frequency} as a metric for empirical computational
complexity~\cite{goldsmith07}.
%Block frequency is a relative metric that represents the number of times a basic
%block executes~\citep{ball94,ball96}.

However, in the context of comparing different optimization sequences,
although block frequency would be able to capture aspects of optimizations that
simplify the control-flow graph (CFG), measuring work at the basic block
resolution would not capture effects of optimisations at the instruction level.
%Because of that, in this thesis, we extend the idea of using basic block
%frequency to measure computational work by also considering the computational
%cost of each basic block.
Because of that, we adopt a work metric that takes into account the
computational cost within each basic block.
The computational cost of a basic block is given by weighting the instructions
that it contains.

@@

We first initialize the edge information $D_{u,v}$, for each edge $(u,v)$ in the CFG.
\[
D_{u,v} \gets 
\begin{cases}
    P_{u,v} & \quad \text{if edge $(u,v)$ has a probe}\\
    0       & \quad \text{otherwise}
\end{cases}   
\]
Afetwards, for each vertex $u$, in post-order of the spanning tree, we apply the following operations:
\[
\Sigma^+_u = \sum_{v\in N^+(u)} D_{u,v}
\]
\[
\Sigma^-_u = \sum_{v\in N^-(u)} D_{v,u}
\]
\[
\forall v\in N^+(u):  D_{u,v} \gets 
\begin{cases}
    D_{u,v} & \quad \text{if $D_{u,v}\neq 0$}\\
    \Sigma^-_u - \Sigma^+_u       & \quad \text{otherwise}
\end{cases}   
\]
\[
\forall v\in N^-(u):  D_{v,u} \gets 
\begin{cases}
    D_{v,u} & \quad \text{if $D_{v,u}\neq 0$}\\
    \Sigma^+_u - \Sigma^-_u       & \quad \text{otherwise}
\end{cases}   
\]

@@

Let $x$ be a symbolic expression, and $\kappa_P(x)$ be the coefficient of the term $P$ in $x$.
In our case, $\kappa_P(x)$ will belong to the set $\{-1,0,1\}$.
Then, for all probes $P$, we have that
\[
w(P) = \sum_u \kappa_P(\Sigma^+_u)w(u)
\]

### 2017-09-12 motivation.tex
minor change
adds formal description of instrumentation

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
    \caption{Comparison between the offline and online {\itercomp}.
             The vertical lines represent the average speedups over
             all the 1,000 inputs.
             }
    \label{fig:motivation-online}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-metric.pdf}
    \caption{Comparison between the offline and online {\itercomp}.
             The vertical lines represent the average speedups over
             all the 1,000 inputs.
             }
    \label{fig:motivation-metric}
\end{figure}

Figure~\ref{fig:motivation-metric} shows a comparison among these three metrics,
namely speedup, execution time, and IPC, and their correlation with the actual
speedup.
Each point in the figure corresponds to each metric averaged over 
a set of inputs, collected during the {\itercomp}.
The vertical lines correspond to the optimization sequence 
with the best score for each one of the metrics.
This figure shows that both execution time and IPC are unsuitable for
online {\itercomp}, as they present no direct correlation with the actual
speedup obtained by the optimization sequence being evaluated.

>>>>

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-online.pdf}
    \caption{Comparison between the offline and online {\itercomp}.
             The vertical lines represent the average speedups over
             all the 1,000 inputs.}
    \label{fig:motivation-online}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{figs/motivation-metric.pdf}
    \caption{Comparison among different performance metrics for performing
             online {\itercomp}.
             The vertical lines represent the best score for each metric.}
    \label{fig:motivation-metric}
\end{figure}

Figure~\ref{fig:motivation-metric} shows a comparison between execution time and IPC,
and their correlation with the actual speedup.
Each point in the figure corresponds to each metric averaged over 
a set of inputs, collected during the {\itercomp}.
The vertical lines correspond to the optimization sequence 
with the best score for each one of the metrics.
This figure shows that both execution time and IPC are unsuitable for
online {\itercomp}, as they present no direct correlation with the actual
speedup obtained by the optimization sequence being evaluated.
